Twist is the key to the gating of mechanosensitive ion channel NOMPC

Version 1

Abstract

Abstract

NOMPC, a tethered mechanosensitive ion channel from the transient receptor potential (TRP) family, can convert mechanical stimuli to ionic electric signals to excite neuron cells. Previous investigations revealed that a pushing force exerted on the linker helix domain or the compression of NOMPC’s ankyrin repeat (AR) domain triggers channel opening. In this study, we explored the direct mechanical cause for NOMPC channel opening as well as the torsional properties of the AR domain using all-atom molecular dynamics simulations. Our results demonstrate that a torque pointing toward the extracellular side, exerted on the amphipathic TRP domain, serves as the direct cause of driving the channel open. The coupling between the compression and twisting of the AR domain ensures that both types of deformations can open the channel. Therefore, we propose a twist-to-open model, facilitated by the compression-twist coupling property of the AR domain, to further elucidate the gating mechanism of the NOMPC channel.

Introduction

Mechanosensitive (MS) ion channels represent pivotal constituents in mechanosensory transduction, playing a fundamental role in the perception of sensory stimuli within living organisms. These mechanosensitive ion channels, coalescing to form mechanical receptors, are strategically positioned within the sensory neuron terminals intricately nestled within the epidermal layer. Their cardinal function lies in the reception and transduction of mechanical signals emanating from the external milieu, ultimately orchestrating their conversion into electrical signals. This intricate process, duly recognized as mechano-electrical transduction (MET), engenders alterations in neuronal excitability, thereby facilitating the transmission of vital information to higher-order neural centers. 1

Among these, NOMPC, a transient receptor potential (TRP) superfamily MS channel identified in C. elegans, Drosophila, and zebrafish, assumes a central role in various mechanosensing-related activities, encompassing soft touch perception, auditory function, and locomotion. 2–6 A salient characteristic distinguishing NOMPC from its counterparts is its possession of a unique intracellular domain comprising twenty-nine ankyrin repeats, stacking one by one into a distinctive spring-like conformation. This exceptional architecture, which was proposed to be associated with microtubules within cellular contexts, designates NOMPC as the only known standalone tether-gating MS channel. 7 Consequently, in contrast to the extensively discussed force-from-lipid MS ion channels, 8,9 an in-depth investigation into the specific molecular mechanism of NOMPC gating stands poised to offer a paradigmatic insight into the realm of the tethered MS ion channels.

In previous studies, it was proposed that compression of the intracellular ankyrin repeat (AR) domain or a pushing force exerted on the AR domain from the intracellular side, could open the channel. 10,11 In the meantime, the TRP domain, widely recognized as crucial in the gating process of TRP channels, 12 was observed to rotate clockwise (CW, looking from the intracellular side) under the torque generated by the pushing force. 11 Therefore, it was still ambiguous whether the compressive force or the twisting torque was the direct mechanical cause behind channel gating. In this work, we determine the critical force component for the gating of NOMPC by using all-atom molecular dynamics (MD) simulations. Our results suggest that the twisting force is the key to the channel gating, while the compression-twist coupling of the AR domain ensures that a compression of the AR domain can activate the channel as well.

Results

A membrane-parallel torsion force was observed on the TRP domain during the push-to-open gating process

For a comprehensive exploration of the atomistic gating details through molecular dynamics (MD) simulations with an affordable computational cost, we employed a divide-and-conquer strategy as depicted in Fig. 1(a). Essentially, the membrane-protein simulation system was divided into two subsystems. System I was composed of the transmembrane (TM) region, intracellular linker helices (LH) domain, and ankyrin repeat 29 (AR29). Within the TM region, there exist six transmembrane helices (S1-S6) in each chain. Four S6 helices of each chain collectively constitute the central pore of NOMPC, and the TRP domain is sandwiched between the TM domain and the LH domain. System II included the LH domain and twenty-nine ankyrin repeat (AR) units. The four spring-like AR chains of the tetrameric protein were organized into a bundle structure, thought to be a key element in mechanical perception. 13–15

Force distribution during the push-to-open gating process of NOMPC. (a) The simulation systems of NOMPC. The NOMPC molecule was divided into two subsystems, denoted by the orange and purple rectangular boxes, for molecular dynamics (MD) simulations. The interface between the Linker helices (LH) and the TRP domains was shown in a circular inset (bottom view). (b) The membrane-parallel (purple line) and membrane-normal (yellow line) components of the net forces on the TRP domain exerted by the LH domain. The error bars denote the standard deviations from two compressive trajectories (CI-1/2) and two free trajectories (FI-1/2). (c) The net forces on the TRP domain exerted by the LH domain. The grey quiver on each residue showed the net force exerted by the LH domain. The black showed the resultant force on the residues of the TRP domain. Two major points of force application, E1571 and R1581, were marked. ⊗ represented the inward force component. The data were averaged for the four subunits by rotationally symmetrized around the protein center.

Building upon the precedent push-to-open model elucidated by Wang et al., 11 we performed steered molecular dynamics (SMD) simulations for system I. In this context, a pushing/compressive force was applied to the ankyrin repeat 29 (AR29), pointing from the intracellular to the extracellular side (trajectories: CI-1/2). Consistent with our previous findings, the transmembrane central pore of NOMPC exhibited a discernible dilation, accompanied by the upward tilt and CW rotation of the TRP domain (Fig. S1). The upward tilt was probably caused by the membrane-normal component of the applied mechanical force, while the CW rotation was generated by the membrane-parallel component that led to a torque pointing to the extracellular side, raising the question that which of these two components was the direct driving force for the channel opening.

To address this pivotal question, we quantitatively analyzed the force distribution in the above simulations CI-1/2, so that we could decompose the force into the membrane-normal and membrane-parallel components to investigate which one could open the channel. In particular, we used the force distribution analysis (FDA) method 16 to quantify the interactions between pairwise residues along the force transmission pathway within system I (Fig. 1(a)). According to our simulation condition and trajectory analyses, the force transmission initiated from the AR29 domain, traversed through the LH domain, and then proceeded to influence the geometrically adjacent yet non-covalently bonded TRP domain, ultimately guiding the S6 domain and instigating the pore opening. Our analysis showed that the strongest interactions between these domains were predominantly located between a few pairs of residues at the interfaces (Fig. S2), including A1118-N1148, Y1109-D1142 on the AR29-LH interface, and K1244-E1571, D1236-R1581, L1256-S1577 on the LH-TRP interface. Their proximity in space made their Coulomb and Lennard-Jones interactions stronger upon the conformational changes. This feature facilitated the stability of hydrogen bonds, particularly the hydrogen bonds on K1244-E1571, D1236-R1581, and L1256-S1577, with occupancy of 57%, 71%, and 24%, respectively. This was consistent with the hydrogen bond pairs identified in previous studies when a pushing force was applied. 11

To provide a clear depiction of the net mechanical force distribution stemming from the externally applied mechanical force, we deducted the FDA results of the force-free simulation trajectories (trajectory: FI-1/2) from the simulation trajectories with pushing/compressive force (trajectory: CI-1/2), and defined the residual force distribution as the net-FDA results (Fig. S3). Analyzing the net mechanical force distribution between the LH and TRP domains using the net-FDA method, we found that, in addition to the membrane-normal force component, the total net mechanical force transmitted from the LH domain to the TRP domain also exhibited a membrane-parallel torsion force component, generating a torque pointing towards the extracellular direction perpendicular to the membrane (Fig. 1(b)-(c)). Moreover, the transmitted force was primarily concentrated on the E1571 and R1581 residues of the TRP domain (Fig. 1(c)).

In brief, the FDA results at the interface between the transmembrane and intracellular domains elucidated a discernible pathway of force transfer, mostly through a few residues located at the AR-LH-TRP interface. Both membrane-normal and membrane-parallel force components were observed on the TRP domain, which provided the basis to further investigate which component is the key factor in channel gating. As previous studies showed that a CW rotation of the TRP domain may be related to the channel opening, it was of particular interest to study the effect of the torsion force on the channel structure.

The membrane-parallel torsion force can open NOMPC, but not the membrane-normal pushing force on the TRP domain

Motivated by the above net-FDA results, we aimed to separate the contributions of the membrane-normal and membrane-parallel force components to the gating process and pinpoint whether one or both of these components played an indispensable role in gating. To achieve this, we needed to determine the point of application, the magnitude, and the direction of the exerted mechanical force. As the TRP domain is sandwiched between the AR and TM domains and its conformational change was proposed to be directly related to the channel gating, it is a natural choice to select the TRP domain to apply the decomposed forces in the MD simulations. Notably, our FDA in the last section showed that the predominant forces propagated from the LH domain to the TRP domain were distributed on two key residues, E1571 and R1581 (Fig. 1(b)-(c)). Therefore, we devised simulations to individually apply membrane-normal and membrane-parallel forces on these two residues of the TRP domain, and the magnitude and direction of the exerted forces were equivalent to those derived from the aforementioned net-FDA results (Fig. 1(b)-(c), Fig. 2(a)-(b)). As the force transmission occurred mainly via side-chain interactions along the interface between the LH and TRP domains, we selected the side chains of the E1571 and R1581 as the point of application of the exerted forces.

A membrane-parallel torsion force can open NOMPC. (a-b) The side view (a) and bottom view (b) of the net-FDA of the E1571 and R1581 on the TRP domain exerted from the LH domain, which was decomposed into the membrane-parallel and membrane-normal components. ⊗ represented membrane-normal forces pointing into the membrane. (c-d) The transmembrane pore size evolution in the simulations with membrane-parallel forces (c) or membrane-normal forces (d), 3 repeats for each condition (trajectories: MPI-1/2/3, MNI-1/2/3, see Methods).

Our simulations revealed that the membrane-parallel torsion force alone was sufficient to open the gate in all three MD simulation repeats (trajectories: MPI-1/2/3, Fig. 2(c)), whereas the membrane-normal pushing component did not trigger pore opening within the simulation time (trajectories: MNI-1/2/3, Fig. 2(d)). According to the pore size evolution in the trajectories MPI-1/2/3 (Fig. 2(c)), it became evident that the most constricted region along the pore, the gate situated around the hydrophobic I1554, exhibited an initial radius measuring less than 1.0 Å, as calculated by the HOLE program. 17 This configuration represented a closed architecture. Under the influence of the membrane-parallel torsion force, the radius of lower constriction was significantly dilated to above 2.0 Å after 600 ns. This dilation signified that water molecules could smoothly pass through the channel (Fig. S4), a phenomenon often indicative of the first step of channel opening. This is highly similar to our previous observation when a pushing force was applied to the AR domain. 11 In contrast, in the trajectories MN-1/2/3 with the membrane-normal force on the TRP domain, no pore dilation was observed (Fig. 2(d)). In fact, the channel appeared to become even more constricted at the upper constriction region (filter), as depicted in Fig. 2(d). This indicated that the membrane-normal force component on the TRP domain was probably not able to open the channel, at least no such sign was observed within our simulation time.

Thus, our investigations substantiated that a torque on the TRP domain, pointing to the extra-cellular side, was the key to driving the channel open. On the basis of the original push-to-open or compress-to-open model, here we provide further details on the force analysis and channel opening, arguing that it is the membrane-parallel torsion force that acts as the direct mechanical cause to open the channel. It should be noted though, this does not mean the push-to-open or compress-to-open model was wrong, as our simulations showed that a pushing force on the LH domain can generate a torsion force on the TRP domain, 11 so does the compression of the AR domain. 10 Apparently, the unique structure of NOMPC ensures that a straightforward pushing force or compression of the AR spring can spontaneously generate a torsion force to open the channel. Therefore, the torsional mechanical properties of the AR domain became more intriguing.

The torsional mechanical properties of the AR domain and force transmission pathway

To study the torsion property of the AR spring, we performed all-atom steered molecular dynamics (SMD) simulations for system II, in which we enforced a gradual clockwise (CW) rotation to the lower end of the AR domain (AR1-AR8) at a rate of 0.1 ° ns−1 along the membrane-normal axis of NOMPC while keeping the upper end (LH domain) position-restrained (Fig. 3(a), trajectory: RII). From the trajectory, we extracted structural snapshots at an interval of 2.5° during the twisting process and relaxed these structures for 100 ns with both ends position-restrained. This protocol was employed to further relax the twisted AR spring to improve the accuracy of the calculated torsional mechanical parameters (Fig. S5). Please refer to the Methods section for more details.

Torsional mechanical properties of the AR bundle. (a) The sketch of the steered molecular dynamics simulation. The dotted arrows revealed the direction of twisting. (b) The relationships between the torque on AR1-8 (left Y-axis), the length of the AR domain (right Y-axis), and the rotational angle of AR1-8. This was calculated from the rotating trajectories RII (0-4).

With the above simulation data, we calculated the torsional coefficient of the AR bundle, using the equation: where M represented the torque applied to both ends, c denoted the torsion coefficient of the AR domain bundle, and ϕ was the average torsion angle of AR1-8 with respect to its initial position. As shown by the green plots in Fig. 3(b), the fitted torsion coefficient turned out to be (2.40 ± 0.06) × 103 kJ mol−1 rad−1. Such a torsion coefficient roughly corresponds to a similarlysized cylindrical volume of uniform polyethylene or rubber material. 18

Interestingly, we observed a gradual increase in the compressive deformation of the AR spring as the torsional angle increased in our simulations (Fig. 3(b)). This implies a coupling between the twist and compression of the AR domain. To quantitatively describe this compression-twist coupling property, we defined the compression-twist coupling coefficient , which represented the compressive deformation induced by a unit torsional deformation. For the AR8-AR29 of NOMPC, the fitted value of the compression-twist coupling coefficient ktc was (1.67 ± 0.14) nm rad−1 within 10 degree torsion. Indeed, it was also noticed that a torque could be generated during the application of compressive forces in previous investigations. 10,11 In light of these findings, we believe that the unique structure of the AR and LH domains bolster NOMPC’s robustness in the sensation of external stimuli, i.e., the compression-twist coupling property of the AR bundle ensures that both the twist and compression of the intracellular domain can generate a torque to rotate the TRP domain, facilitating the opening of the pore.

To pinpoint the force transmission pathway along the AR spring while it is twisted, we utilized the same net-FDA method as above to analyze the force distribution during the twisting process of the trajectory RII. The net-FDA results of each (i + 1)th AR unit exerted from the adjacent ith AR unit unveiled a spiral force transmission route, rooted in the helical structure inherent to the AR spring (Fig. 4(a)). Strong pairwise interactions predominantly resided in bonded residue pairs, while only a small fraction was associated with non-bonded residue pairs (Fig. 4(b)). Clearly, these bonded residue pairs, which are integral to the spiral backbone of the AR spring, were critical for transferring external forces. The non-bonded residue pairs were primarily situated at the adjacent turns within the AR α-helices and the interfaces between AR units. To clearly illustrate the details, we have magnified the local structure of AR13-15 (Fig. S6). As depicted in Fig. 4(c)-left, a representative non-bonded net-FDA of the local region of AR13-15 was shown. The data around the diagonal line represented the pairwise residues located at the adjacent turns between α-helices, while the wing-shaped data positioned further away from the diagonal indicated pairwise residues situated at the interfaces between AR units. Apparently, although the bonded residue pair interactions are predominant, the non-bonded interactions, both within the same AR unit and across different AR units, are important for the force transmission too.

Force transmission along the AR domain. (a) The quiver plot of net-FDA (XY planar) of each (i +1)th AR unit exerted from the ith AR unit showed the force transmission from the intracellular terminal to the transmembrane side of the AR domain viewing from AR1 toward AR29. ⊙represented the outward Z-component force. (b) The distribution of the pairwise residue-residue net-FDA within AR9 to AR29 (weak interaction under 5 pN neglected). The inset showed local information over 70 pN. (c) The representative interaction heatmaps (AR13-AR15) showed the magnitude of non-bonded pairwise residue-residue net-FDA (left) and hydrogen bonds occupancy (right). (d) The Pearson correlation coefficient (PCC) between net-FDA and hbonds in AR units. Inset showed the mean value and standard deviation of Intra-AR (diagonal) and Inter-AR (sub-diagonal) PCCs.

Furthermore, we noticed that there is a correlation between the distribution of non-bonded net-FDA and the existence/strength of hydrogen bonds, as demonstrated by similar patterns in Fig. 4(c). For clarity, we divided these two 737× 737 matrices (ResID of AR9-AR29: 400-1136) into 21× 21 subgroups (grouped by AR unit). Then, we calculated the Pearson correlation coefficient (PCC) between the non-bonded net-FDA and the occupancy of hydrogen bonds of each subgroup, resulting in a 21× 21 PCC matrix (Fig. 4(d)). Relatively high PCCs in the sub-diagonal grids of Fig. 4(d) (inter-AR), and even higher PCCs in the diagonal grids (intra-AR), are observed (inset of Fig. 4(d)). This underscored the crucial role of the hydrogen bonding network in facilitating force transmission. Therefore, we concluded that the force transmission route primarily consisted of bonded residues that formed the spiral backbone of the AR chains, with supplementary support from the hydrogen bonding network formed by adjacent residues in the same or neighboring AR unit.

Discussion

The unique structure of NOMPC and its fascinating AR spring have sparked great curiosity. Building upon the push-to-open and compress-to-open mechanism, here we propose a twist-to-open model, as illustrated in Fig. 5. At the core of this model was the idea that a twisting force on the TRP domain is the key to the opening of the pore. This is not to say the compress-to-open or push-to-open models are wrong, but that a compression/pushing on the AR domain will lead to a twist force on the TRP domain, and this twist force is the direct cause of the pore opening. The spiral structure of each AR chain, as well as the hydrogen bonding network formed among spatially adjacent residues, enables the AR bundle to convert the compression deformation to a twist deformation, thus exerting a membrane-parallel torsional force on the LH domain. In addition, the structure around the LH domain can also convert a membrane-normal pushing force to a membrane-parallel torsional force, as shown in our previous study. 11 Consequently, the compression of the AR domain, or a pushing force from the intracellular side, can be converted to a torque on the TRP domain, which points to the extracellular side. We believe this torque is the direct mechanical cause that drives the TRP domain to rotate clockwise (looking from the intracellular side) to open the channel.

A schematic diagram of the twist-to-open model. The side view (left) and bottom view (right) of the twist-to-open model. Both the compression and CW twisting of the AR domain can generate a membrane-parallel twisting force on the TRP domain, which is the key component for gating.

The highly conserved TRP domain serves as a structural characteristic of the TRP protein superfamily. In our simulations of NOMPC, we observed a clockwise (CW) rotation of the TRP domain during the gating process of the proteins. Interestingly, such rotation was also observed in other TRP channels. For example, the recently resolved cryo-EM structures of the thermo/ligand-gated mm-TRPM8 (PDB: 8E4P, 8E4N, 8E4M, 8E4L; 19), thermo/ligand-gated rat-TRPV1 (PDB: 7L2H, 7L2M, 7L2N, 7L2L, 7MZ5; 20), and ligand-gated homo-TRPV4 (PDB: 6BBJ, 7AA5; 21) showed similar gating motion of the TRP domain. Based on these findings, we believe that the rotation of the TRP domain could be a common feature in the gating mechanism of these TRP proteins, and the unique structure and compression-twist coupling property of the intracellular domain of NOMPC ensure its robustness in the perception of multiple forms of mechanical stimuli: not only twist, but also compression.

Although we calculated the detailed interactions between residue pairs of NOMPC, it should be noted that, due to the limitations of the FDA method, our force analysis lacked the entropic force, which may also play a significant role in the force transmission along the AR spring. Nonetheless, this won’t affect our twist-to-open model, since the entropic effect was present in our MD simulations. Another limitation of the present study is that the simulation time is very short compared to the physiological timescale. However, we found that the mechanical stimuli are transferred at a fast speed along AR springs in simulations, at around 1.8 nm ps−1. 11 Therefore, the mechanical stimuli were nearly instant for our simulation time, which allowed us to determine the trend or the direction of the conformational changes, even though the observation of the complete gating process is still challenging. In this context, we think that the dilation of the pore under a torsion force on the TRP domain is a piece of strong evidence that twist is the key to the channel opening.

Methods

The all-atom simulation Systems

To set up the molecular dynamics (MD) simulations, we first prepared the simulation systems using the cryo-EM structure of NOMPC (PDB: 5VKQ). 7 CHARMM-GUI was used to generate the configuration and topology of the simulation systems. 22 The parameter files were created for the CHARMM36m force field. 23,24

The system I included the TM domain, the LH domain, and the AR29 of NOMPC. The PPM server was used to reorient the NOMPC structure to ensure that the TM domain of NOMPC was well located in a lipid bilayer. 25 The protein was embedded in a POPC bilayer. Then, we solvated the protein-membrane complex in a water box using the TIP3P water model. Counterions were added to neutralize the system, ensuring physiological ionic strength by adding Na+ and Cl− ions to achieve a concentration of 0.15 M. The details of simulation system I are shown in Table 1.

The system II included the LH domain and the AR domain (AR1-AR29) of NOMPC. The protein structures were immersed in a rectangular box of TIP3 water molecules and 0.15 M NaCl. The details of simulation system II are shown in Table 1.

All-atom MD simulations

All of the MD simulations were performed with GROMACS 5.1.2. 26 The REDUCE program in AMBER was used to add hydrogens to the original PDB files and determine the protonation state of the histidine residues. 27,28

For system I, we used the steepest descent algorithm to achieve energy minimization. Then, following up with a two-stage equilibration, a 0.4 ns NVT equilibration simulation with harmonic restraint was applied to the protein molecule (force constants of 4000 kJ mol−1 nm−2 on the back-bone and 2000 kJ mol−1 nm−2 on the side chains), and a 20 ns NPT equilibration simulation with gradually decreased restraint (from 2000 to 100 kJ mol−1 nm−2 on the backbone and from 1000 to 50 kJ mol−1 nm−2 on the side chains). During the equilibration processes, planar restraints were used to keep the positions of lipid head groups along the membrane-normal direction.

For system II, we used the steepest descent algorithm to achieve energy minimization. Then, also followed by a two-stage equilibration, a 0.2 ns NVT equilibration simulation with harmonic restraint was applied to the protein molecule (force constants of 400 kJ mol−1 nm−2 on the backbone and 40 kJ mol−1 nm−2 on the side chains), and a 10 ns NPT equilibration simulation with harmonic restraint (force constants of 400 kJ mol−1 nm−2 on the backbone and 40 kJ mol−1 nm−2 on the side chains).

The simulation temperature of the system was set to 300 K. The time step was 2 fs. The cubic periodic boundary condition was used during the simulations, and the van der Waals interaction was switched off from 10 to 12 Å. The long-range electrostatic interactions were calculated using the particle mesh Ewald (PME) method. 29

Steered MD simulations

Steered molecular dynamics (SMD) simulations were utilized to probe the mechanical gating mechanisms of the ion channels. SMD was performed by attaching a virtual spring to a specific group of atoms, and a constant velocity or constant force was applied to the spring to pull the molecule along a predefined reaction coordinate. The force exerted on the molecule was recorded throughout the simulation. To ensure reproducibility, multiple independent SMD simulations were performed in this research.

The SMD simulation detail of the compressive trajectories of system I (CI-1/2) and free trajectories of system I (FI-1/2) can be found in previous work. 11

For trajectories MPI-1/2/3 (membrane-parallel force component on system I Repeat-1/2/3) and MNI-1/2/3 (membrane-normal force component on system I Repeat-1/2/3) of system I (Fig. 2(c)-(d)), we separately exerted membrane-parallel (MPI-1/2/3) or membrane-normal (MNI-1/2/3) external force on the side chain of E1571 and R1581, as shown in Table 2. At the first 200 ns, we applied an increasing force by using the constant-rate pulling method. Then we applied the calculated constant force by using the constant-force pulling method in the last 600 ns. To prevent the protein from rotation as a whole, we used the enforced rotation method, 30 setting the rotation speed as zero for the S1-S4 helices in the transmembrane region of the protein. Distance restraints were applied between the acceptors and donors on the side chains of residue pairs 1564-1568 and 1567-1571, to maintain the hydrogen bonds and prevent structural damage in the α-helix under the applied mechanical force.

For trajectory RII (Rotation of system II), we performed a three-step simulation to better study the torsional mechanical properties of the AR domain, as shown in Table 3.

In the first step, we aimed to relax the structure by fixing the relative rotational angle between both ends along the rotational axis, which is the Z axis of system II. We choose to manipulate AR1-AR8 as a whole, considering the fact that the cryo-EM structure of AR1-AR7 was not resolved but modeled. Therefore, it was prudent to exert and transmit torque starting from the more reliably resolved structure, i.e., from AR8 upward, while preserving the integrity of the entire protein structure. With the help of the enforced rotation function of GROMACS, 30 we used the pm rotational potential on the bottom terminal of the AR domain (AR1-AR8) and the up terminal (LH domain) with a zero angular velocity for 20 ns. The pm rotational potential can constrain the group of atoms to rotate within the rotational plane as required, while allowing free movement in the direction perpendicular to the rotational plane. This means that the entire AR bundle will be subjected to torsion, while being able to stretch and compress freely.

In the second step, we clockwise rotated the AR1-8 along the Z axis at an angular velocity of 0.1 ° ns−1, while keeping the LH domain still (Fig. 3(a)) with pm rotational potential. The difference in angular velocity between the bottom and up terminal accumulated a specific rotation angle within 10° (Fig. S5, trajectory: RII-0).

In the third step, when rotating at every 2.5° in trajectory RII-0, we extracted structural snapshots and held both ends of the AR domain for 100 ns to get a better equilibrated structure (Fig. S5, trajectory: RII-1 for the 2.5°-rotational angle, RII-2 for 5°, RII-3 for 7.5°, and RII-4 for 10°), and the readouts of torques on both ends gradually increased with opposite directions. Specifically, we used the trajectory RII-1/2/3/4 to calculate the torsional mechanical properties of the AR bundle and the trajectory RII-0 to study the force transmission pathway along the AR chain. When the torsional angles of the AR bundle are 2.5i° (i = 1, 2, 3, 4), we analyze the 100 ns-long RII-i trajectories. For the torques at both ends of the AR, we collect torque data on the LH domain every 2 ps. Under equilibrium conditions, the torques at the top and bottom of the AR bundle are equal in magnitude but opposite in direction. To assess the length change of the AR bundle, we measure the distance between the centroids of the LH domain and AR8 unit.

Force distribution analysis

Force distribution analysis (FDA) focuses on computational efficiency and low-memory usage and makes possible time series analysis of pairwise forces variation in long MD simulations and for large molecular systems. 16 It can compute atomic pairwise forces for all types of bonded interactions as well as Coulomb and Lennard-Jones potential by rerunning the trajectories. By summing up all the atomic pairwise forces within the two residues of interest, it can also calculate the pairwise force of the interaction between residues ri and rj, which is labelled as . Similarly, if we focus on the interaction between two domains, we can sum up all the pairwise forces of the residues from these two domains.

In the calculation, we made an input file which contained all the parameters that FDA needed to run. We set onepair = detailed and type = all so that all pairwise forces of different types between the same pair of residues are stored separately. The option residuebased was set as pairwise_forces_vector so that the residue-based pairwise forces were written to output files as vector values. The forces for every frame were written by setting time_averages_period = 0.

Then, forces sampled by MD simulations under two conditions (with and without perturbation) were compared. We analyzed the interactions between residues in the trajectories with the external force (CI-1/2), labelled as , and without the external force (FI-1/2), labelled as , respectively. Then the average net force distribution induced by the compressive external force, which was named net-FDA, , was obtained by subtracting the FDA result of trajectories without the external force from the FDA result of trajectories with the external force,

In the end, we averaged the vector net-FDA results of the four subunits by rotationally symmetrized around the protein center.

eLife Assessment

This valuable study uses steered molecular dynamics simulations to interrogate force transmission in the mechanosensitive NOMPC channel, which plays roles including soft-touch perception, auditory function, and locomotion. The finding that the ankyrin spring transmits force through torsional rather than compression forces may help understand the entire TRP channel family. The evidence is, however, considered to be still incomplete. It could be strengthened by testing how the channel responds to different twisting and compressional force magnitudes over longer simulation times to see a full gating motion, or to prove that the partial or initial motion observed relates to physiological gating. Experimental validation of reduced mechano-sensitivity through mutagenesis of proposed ankyrin/TRP domain coupling interactions would be best to enhance the manuscript.

Reviewer #1 (Public review):

Summary:

This manuscript uses molecular dynamics simulations to understand how forces felt by the intracellular domain are coupled to the opening of the mechanosensitive ion channel NOMPC. The concept is interesting - as the only clearly defined example of an ion channel that opens due to forces on a tethered domain, the mechanism by which this occurs is yet to be fully elucidated. The main finding is that twisting of the transmembrane portion of the protein - specifically via the TRP domain that is conserved within the broad family of channels- is required to open the pore. That this could be a common mechanism utilised by a wide range of channels in the family, not just mechanically gated ones, makes the result significant. It is intriguing to consider how different activating stimuli can produce a similar activating motion within this family. However, the support for the finding can be strengthened as the authors cannot yet exclude that other forces could open the channel if given longer or at different magnitudes. In addition, they do not see the full opening of the channel, only an initial dilation. Even if we accept that twist is essential for this, it may be that it is not sufficient for full opening, and other stimuli are required.

Strengths:

Demonstrating that rotation of the TRP domain is the essential requirement for channel opening would have significant implications for other members of this channel family.

Weaknesses:

The manuscript centres around 3 main computational experiments. In the first, a compression force is applied on a truncated intracellular domain and it is shown that this creates both a membrane normal (compression) and membrane parallel (twisting) force on the TRP domain. This is a point that was demonstrated in the authors' prior eLife paper - so the point here is to quantify these forces for the second experiment.

The second experiment is the most important in the manuscript. In this, forces are applied directly to two residues on the TRP domain with either a membrane normal (compression) or membrane parallel (twisting) direction, with the magnitude and directions chosen to match that found in the first experiment. Only the twisting force is seen to widen the pore in the triplicate simulations, suggesting that twisting, but not compression can open the pore. This result is intriguing and there appears to be a significant difference between the dilation of pore with the two force directions. However, there are two caveats to this conclusion. Firstly, is the magnitude of the forces - the twist force is larger than the applied normal force to match the result of experiment 1. However, it is possible that compression could also open the pore at the same magnitude or if given longer. It may be that twist acts faster or more easily, but I feel it is not yet possible to say it is the key and exclude the possibility that compression could do something similar. I also note that when force was applied to the AR domain in experiment 1, the pore widened more quickly than with the twisting force alone, suggesting that compression is doing something to assist with opening. Given that the forces are likely to be smaller in physiological conditions it could still be critical to have both twist and compression present. As this is the central aspect of the study, I believe that examining how the channel responds to different force magnitudes could strengthen the conclusions and recommend additional simulations be done to examine this.

The second important consideration is that the study never sees a full pore opening, but rather a widening that is less than that seen in open state structures of other TRP channels and insufficient for rapid ion currents. This is something the authors acknowledge in their prior manuscript in eLife 2021. While this may simply be due to the limited timescale of the simulations, it needs to be clearly stated as a caveat to the conclusions. Twist may be the key to getting this dilation, but we don't know if it is the key to full pore opening. To demonstrate that the observed dilation is a first step in pore opening, then a structural comparison to open-state TRP channels would be beneficial to provide evidence that this motion is along the expected pathway of channel gating.

Experiment three considers the intracellular domain and determines the link between compression and twisting of the intracellular AR domain. In this case, the end of the domain is twisted and it is shown that the domain compresses, the converse to the similar study previously done by the authors in which compression of the domain was shown to generate torque. While some additional analysis is provided on the inter-residue links that help generate this, this is less significant than the critical second experiment.

Reviewer #2 (Public review):

This study uses all-atom MD simulation to explore the mechanics of channel opening for the NOMPC mechanosensitive channel. Previously the authors used MD to show that external forces directed along the long axis of the protein (normal to the membrane) result in AR domain compression and channel opening. This force causes two changes to the key TRP domains adjacent to the channel gate: 1) a compressive force pushes the TRP domain along the membrane normal, while 2) a twisting torque induces a clock-wise rotation on the TRP domain helix when viewing the bottom of the channel from the cytoplasm. Here, the authors wanted to understand which of those two changes is responsible for increasing the inner pore radius, and they show that it is the torque. The simulations in Figure 2 probe this question with different forces, and we can see the pore open with parallel forces in the membrane, but not with the membrane-normal forces. I believe this result as it is reproducible, the timescales are reaching 1 microsecond, and the gate is clearly increasing diameter to about 4 Å. This seems to be the most important finding in the paper, but the impact is limited since the authors already show how forces lead to channel opening, and this is further teasing apart the forces and motions that are actually the ones that cause the opening.

Reviewer #3 (Public review):

Summary:

This manuscript by Duan and Song interrogates the gating mechanisms and specifically force transmission in mechanosensitive NOMPC channels using steered molecular dynamics simulations. They propose that the ankyrin spring can transmit force to the gate through torsional forces adding molecular detail to the force transduction pathways in this channel.

3. Constant velocity or constant force

For the SMD the authors write "and a constant velocity or constant force". It's unclear from this reviewer's perspective what is used to generate the simulation data.

Strengths:

Detailed, rigorous simulations coupled with a novel model for force transduction.

Weaknesses:

Experimental validation of reduced mechanosensitivity through mutagenesis of proposed ankyrin/TRP domain coupling interactions would greatly enhance the manuscript. I have some additional questions documented below:

(1) The membrane-parallel torsion force can open NOMPC

How does the TRP domain interact with the S4-S5 linker? In the original structural studies, the coordination of lipids in this region seems important for gating. In this manner does the TRP domain and S4-S5 linker combined act like an amphipathic helix as suggested first for MscL (Bavi et al., 2016 Nature Communications) and later identified in many MS channels (Kefauver et al., 2020 Nature).

(2) Torsional forces on shorter ankyrin repeats of mammalian TRP channels

Is it possible torsional forces applied to the shorter ankyrin repeats of mammalian TRPs may also convey force in a similar manner?

(3) Constant velocity or constant force

For the SMD the authors write "and a constant velocity or constant force". It's unclear from this reviewer's perspective which is used to generate the simulation data.

Author response:

Reviewer #1 (Public review):

Summary:

This manuscript uses molecular dynamics simulations to understand how forces felt by the intracellular domain are coupled to the opening of the mechanosensitive ion channel NOMPC. The concept is interesting - as the only clearly defined example of an ion channel that opens due to forces on a tethered domain, the mechanism by which this occurs is yet to be fully elucidated. The main finding is that twisting of the transmembrane portion of the protein - specifically via the TRP domain that is conserved within the broad family of channels- is required to open the pore. That this could be a common mechanism utilised by a wide range of channels in the family, not just mechanically gated ones, makes the result significant. It is intriguing to consider how different activating stimuli can produce a similar activating motion within this family. However, the support for the finding can be strengthened as the authors cannot yet exclude that other forces could open the channel if given longer or at different magnitudes. In addition, they do not see the full opening of the channel, only an initial dilation. Even if we accept that twist is essential for this, it may be that it is not sufficient for full opening, and other stimuli are required.

Strengths:

Demonstrating that rotation of the TRP domain is the essential requirement for channel opening would have significant implications for other members of this channel family.

Thank you for your positive summary and comments.

Weaknesses:

The manuscript centres around 3 main computational experiments. In the first, a compression force is applied on a truncated intracellular domain and it is shown that this creates both a membrane normal (compression) and membrane parallel (twisting) force on the TRP domain. This is a point that was demonstrated in the authors’ prior eLife paper - so the point here is to quantify these forces for the second experiment.

The second experiment is the most important in the manuscript. In this, forces are applied directly to two residues on the TRP domain with either a membrane normal (compression) or membrane parallel (twisting) direction, with the magnitude and directions chosen to match that found in the first experiment. Only the twisting force is seen to widen the pore in the triplicate simulations, suggesting that twisting, but not compression can open the pore. This result is intriguing and there appears to be a significant difference between the dilation of pore with the two force directions.

However, there are two caveats to this conclusion. Firstly, is the magnitude of the forces - the twist force is larger than the applied normal force to match the result of experiment 1. However, it is possible that compression could also open the pore at the same magnitude or if given longer. It may be that twist acts faster or more easily, but I feel it is not yet possible to say it is the key and exclude the possibility that compression could do something similar.

Thank you for your insightful comment. As you pointed out, the membranenormal pushing forces exerted at residues E1571 and R1581 are approximately onethird and two-thirds, respectively, of the membrane-parallel twisting forces. These magnitudes were derived from a previous simulation (Wang et al., 2021), in which we decomposed the resultant force into its membrane-parallel and membrane-normal components upon applying a compressive force to the intracellular AR end. Our results indicated that, upon reaching the TRP helix, the induced twisting force is indeed greater, which partially reflects actual physiological conditions. Therefore, considering the magnitudes of the resultant forces alone, the twisting force is predominantly greater than the pushing force when the AR domain is subjected to compression.

Then the question became, if forces of the same magnitude are applied in either the membrane-normal or membrane-parallel directions, what would the outcome be? To address this, we conducted additional simulations. Considering the situations discussed above, we applied a smaller membrane-parallel force instead of a larger membranenormal force that may disrupt the integrity of protein and membrane structure. As shown in the new Figure S6, we adjusted the applied membrane-parallel force to either half or one-third of the original value. When we applied half of the force used in the original setup, the channel opened in two out of three trajectories. When applying onethird of the force, the channel opened in one out of three trajectories. Together with our previous results, these findings suggest that if forces of equal magnitude are applied in the membrane-normal and membrane-parallel directions, the membrane-parallel force has a higher probability of inducing channel opening.

Still, one cannot completely exclude the possibility that the pushing force on the TRP helix can open the channel if given a very long time. This becomes unfeasible to examine with MD simulations, so we investigated the likely conformational changes of multiple TRP family proteins upon opening, and found that the TRP rotation is a universal conformational change, while the TRP tilt is much less consistent (Figure 6). These findings gives us more confidence that the twist force plays a more crucial role in channel gating than the pushing force. We have added a new table (Table 1) and a new figure (Figure 6) to present this analysis.

In addition, we did not intend to imply that compression is incapable of contributing to channel opening. In fact, our aim was to highlight that compression can generate both a twisting force and a pushing force, with the twisting force appearing to be the more critical component for facilitating channel opening. We concur that we cannot completely dismiss the possibility that the pushing component may also assist in channel opening. Consequently, we have revised our discussion on pages 4,6 to enhance clarity.

I also note that when force was applied to the AR domain in experiment 1, the pore widened more quickly than with the twisting force alone, suggesting that compression is doing something to assist with opening.

You are correct that the trajectory corresponding to Experiment 1 (Figure S1(b)) indicates pore opening around 300-400 ns, while the trajectory for Experiment 2 (800 ns) shows pore opening around 600 ns. This observation may suggest that the pore opens more rapidly in Experiment 1, assuming that the simulation conditions were identical for both experiments. However, it is important to note that in Experiment 1, an external force was applied to AR29. In contrast, in Experiment 2, the force was applied exclusively to two selected residues on the TRP domain, while other TRP residues also experienced mechanical forces, albeit to a lesser extent. The differing methods of force application in the two experiments complicate the comparison of pore opening speeds under these conditions.

We acknowledge that the compression of the AR spring can facilitate pore opening. This compression generates both a twisting component and a pushing component on the TRP domain. Our simulations and structural analyses of multiple TRP channels suggest that the twisting component plays a predominant role in gating. However, we cannot entirely rule out the possibility that the pushing component may also contribute to this process. We have carefully revised our Result (page 6), Discussion (pages 10–12) and Methods (pages 14–17) sections to enhance clarity.

Given that the forces are likely to be smaller in physiological conditions it could still be critical to have both twist and compression present. As this is the central aspect of the study, I believe that examining how the channel responds to different force magnitudes could strengthen the conclusions and recommend additional simulations be done to examine this.

Thank you for your valuable comments. We agree that the force applied in Experiment 2 is possible to be larger than the physiological conditions. Therefore, we performed additional simulations to investigate the possibility of opening the pore using smaller torsional forces.

As shown in the new Figure S6, we applied half and one-third of the original force and performed three replicate simulations for each condition. With half the force, the pore opened in two out of the three simulations. And with one-third of the applied force, the pore opened in one out of the three replicate simulations. The probability of pore opening within the same simulation time decreased as the applied force was reduced, consistent with our expectations. These new results are provided as supplementary figures (Figure S6) in the revised manuscript.

We anticipate that further reductions in the forces will result in additional delays in the opening process; however, this would lead to prohibitive computational costs. Consequently, we have decided to conclude our analysis at this stage and have discussed this matter on page 6 of the revised manuscript.

The second important consideration is that the study never sees a full pore opening, but rather a widening that is less than that seen in open state structures of other TRP channels and insufficient for rapid ion currents. This is something the authors acknowledge in their prior manuscript in eLife 2021. Although this may simply be due to the limited timescale of the simulations, it needs to be clearly stated as a caveat to the conclusions. Twist may be the key to getting this dilation, but we do not know if it is the key to full pore opening. To demonstrate that the observed dilation is a first step in the opening of pores, a structural comparison to open-state TRP channels would be beneficial in providing evidence that this motion is along the expected pathway of channel gating.

We are grateful for this insightful comment. We acknowledge that our simulations do not capture a fully open state, but rather a dilation that is smaller than the open-state structures of other TRP channels. In our simulations, a pore radius exceeding 2 Å is considered as a partially open state, as this is generally sufficient for the permeation of water molecules or even small cations such as K+ and Na+ However, the passage of larger molecules and ions, such as Ca2+ and clusters of hydrated ions, remains challenging. As you noted, this partial opening may be attributed to the limited timescale of the simulations.

Furthermore, in accordance with your suggestion, we analyzed numerous TRP proteins for which multiple open or intermediate states have been resolved, and we have included a new figure (Figure 6). A clockwise rotation of the TRP domain is observed in the majority of these proteins upon gating. For instance, in the case of RnTRPV1, our analysis revealed that during TRPV1 activation, when different ligands are bound (RTX, DkTX), the pore undergoes gradual dilation, which involves a progressive clockwise rotation of the TRP domain. This analysis provides evidence that the observed motion aligns with expected gating transitions, supporting the notion that twist-induced TRP rotation and pore dilation may represent an initial step in the pore opening process.

Nonetheless, we concur that further studies, including extended simulations, which are currently unfeasible, or experimental validation, will be necessary to ascertain whether our proposed mechanism is adequate for the complete opening of the pore. We have carefully discussed this on pages 10–12.

Experiment three considers the intracellular domain and determines the link between compression and twisting of the intracellular AR domain. In this case, the end of the domain is twisted and it is shown that the domain compresses, the converse to the similar study previously done by the authors in which compression of the domain was shown to generate torque. While some additional analysis is provided on the inter-residue links that help generate this, this is less significant than the critical second experiment.

Although experiment three is less significant in revealing the underlying gating mechanism, it provides quantitative measurements of the mechanical properties of the intriguing AR spring structure, which are currently challenging to obtain experimentally. These provide computational predictions for future experiments to validate.

Reviewer #2 (Public review):

This study uses all-atom MD simulation to explore the mechanics of channel opening for the NOMPC mechanosensitive channel. Previously the authors used MD to show that external forces directed along the long axis of the protein (normal to the membrane) result in AR domain compression and channel opening. This force causes two changes to the key TRP domains adjacent to the channel gate: 1) a compressive force pushes the TRP domain along the membrane normal, while 2) a twisting torque induces a clock-wise rotation on the TRP domain helix when viewing the bottom of the channel from the cytoplasm. Here, the authors wanted to understand which of those two changes is responsible for increasing the inner pore radius, and they show that it is the torque. The simulations in Figure 2 probe this question with different forces, and we can see the pore open with parallel forces in the membrane, but not with the membrane-normal forces. I believe this result as it is reproducible, the timescales are reaching 1 microsecond, and the gate is clearly increasing diameter to about 4 Å. This seems to be the most important finding in the paper, but the impact is limited since the authors already show how forces lead to channel opening, and this is further teasing apart the forces and motions that are actually the ones that cause the opening.

Thank you for your insightful comments. We appreciate your recognition of our key finding that torque is responsible for increasing the inner pore radius. Indeed, our simulations illustrated in Figure 2 systematically explore the effects of different forces on pore opening. These results demonstrate that membrane-parallel forces are effective, while membrane-normal forces are not within the simulation time. We acknowledge that this study builds upon previous findings regarding force-induced channel opening. However, we believe that further decomposition of the specific forces and motions responsible for this process provides valuable mechanistic insights. By distinguishing the role of torque from the membrane-normal forces of the TRP helix, which is highly conserved across the TRP channel family, our work contributes to a more precise understanding of TRP channel gating. Moreover, in the revised manuscript, we conducted a systematic analysis of the structures of TRP family proteins and discovered that the clockwise rotation of the TRP domain is likely a universal gating mechanism among the TRP family, which significantly enhances and strengthens our original findings (Figure 6).

Reviewer #3 (Public review):

Summary:

This manuscript by Duan and Song interrogates the gating mechanisms and specifically force transmission in mechanosensitive NOMPC channels using steered molecular dynamics simulations. They propose that the ankyrin spring can transmit force to the gate through torsional forces adding molecular detail to the force transduction pathways in this channel.

Strengths:

Detailed, rigorous simulations coupled with a novel model for force transduction.

Thank you for your positive comments.

Weaknesses:

Experimental validation of reduced mechanosensitivity through mutagenesis of proposed ankyrin/TRP domain coupling interactions would greatly enhance the manuscript. I have some additional questions documented below:

We attempted to measure the mechanical properties of the AR domain and conduct mutagenesis experiments in collaboration with Prof. Jie Yan’s laboratory at the Mechanobiology Institute, National University of Singapore; however, this proved to be a significant challenge at this time. Given the urgency of the publication, we have decided to first publish the computational results and reserve further experimental studies for future investigations.

(1) The membrane-parallel torsion force can open NOMPC

How does the TRP domain interact with the S4-S5 linker? In the original structural studies, the coordination of lipids in this region seems important for gating. In this manner does the TRP domain and S4-S5 linker combined act like an amphipathic helix as suggested first for MscL (Bavi et al., 2016 Nature Communications) and later identified in many MS channels (Kefauver et al., 2020 Nature).

In our analysis of the compression trajectories (trajectory: CI-1, Figure S4), we identified stable interactions between the TRP domain and the S4-S5 linker. These interactions primarily involve the residues S1421 and F1422 of the S4-S5 linker, as indicated by the large pink data points in Figure S4. Therefore, we agree that the TRP helix and the S4–S5 linker can be considered an amphipathic helical unit, analogous to the amphipathic helix observed in MscL and other mechanosensitive channels. Moreover, the pocket adjacent to the S4-S5 linker has been recognized as a binding site for small molecules in other ligand-activated TRP channels, such as the vanilloid-binding TRPV1. We hypothesize that this unit is likely to play a critical role in the polymodal gating of the TRP channel family, including ligand-induced activation. In the revised manuscript, we have included an analysis of the interaction between the TRP domain and the transmembrane (TM) domain on page 4 (Figure S4), and we have briefly discussed its implications on pages 10 and 12.

(2) Torsional forces on shorter ankyrin repeats of mammalian TRP channels

Is it possible torsional forces applied to the shorter ankyrin repeats of mammalian TRPs may also convey force in a similar manner?

This is an intriguing question.

To answer your question, we studied the full-length squirrel TRPV1 (PDB: 7LQY, Nadezhdin et al. (2021)) using all-atom steered MD simulations. We applied pushing or torsional forces to the intracellular AR1-2 region of TRPV1, separately (Figure S10(a)). Similar to NOMPC, rotation of the TRP domain was observed under both types of mechanical stimulation (Figure S10(b-e)). The conformational change induced by the torsional force on the TRP domain resembles the change observed in NOMPC. This suggests that a torsional force applied to the shorter ankyrin repeats of mammalian TRPs may yield similar effects on channel gating. However, given that these ankyrin repeats do not act like tether elements, the implications of these results in the context of biological functions remain unclear. Additionally, in NOMPC, the AR domain is connected to the TRP domain through a linker helix (LH) domain, composed of multiple stacked helices that form a relatively compact structure (Figure 1(a)). In contrast, TRPV1 does not possess a similarly compact LH domain connecting the AR domain to the TRP domain (Figure S10(a)). These structural differences render our conclusions regarding NOMPC not directly applicable to TRPV1. We have included an additional discussion about this on page 12 (Figure S10).

(3) Constant velocity or constant force

For the SMD the authors write "and a constant velocity or constant force". It’s unclear from this reviewer’s perspective which is used to generate the simulation data.

Thank you for pointing out this ambiguity. In our simulations, we first applied constant-velocity pulling to achieve specific force magnitudes, followed by constantforce pulling. This protocol allowed us to initiate the motion of the protein in a controlled manner and observe the response of the system under sustained forces. We have now clarified this in the revised Methods section.

Reviewer #1 (Recommendations for the authors):

The language in the paper requires some editing - particularly in the introduction. For example, what is meant by ion channels ’coalescing to form mechanical receptors’? Are the authors implying it requires multiple channels to form a receptor? It is stated that mechanically gated ion channels are only found in nerve endings when in fact they are found in almost every cell type. Another example is the statement ’In the meantime’ the TRP domain was observed to rotate when this observation came prior to the others mentioned before. While these sound like minor edits, they significantly change the meaning of the introduction. I recommend careful editing of the manuscript to avoid accidental inaccuracies like this.

Thank you for your feedback on the clarity and accuracy of the introduction. We have carefully revised the manuscript, particularly the abstract and instroduction sections, to address these concerns:

(1) We have reworded the original sentence ’These mechanosensitive ion channels, coalescing to form mechanical receptors, are strategically positioned within the sensory neuron terminals intricately nestled within the epidermal layer.’ into ’In both vertebrates and invertebrates, mechanosensitive ion channels are widely expressed in peripheral sensory neurons located near or within the surface tissues responsible for detecting mechanical stimuli.’

(2) We have replaced the phrase "In the meantime" with "Interestingly" to introduce the conformational change of the TRP domain that we believe is crucial.

(3) We have carefully reviewed the entire manuscript and used a language editing tool, Writefull integrated within Overleaf, to proof-check the language problems.

Reviewer #2 (Recommendations for the authors):

How do the energy values in Figure 3b, compare with the continuum energy values reported by Argudo et al. JGP (2019)? I wonder what value the authors would get with a new replicate run slower - say 200 ns total aggregate simulation? This would probe the convergence of this energy value. It seems important to determine whether the loading velocity of the experiments performed here with the steered MD is slow enough to allow the protein to relax and adopt lower energy configurations during the transition. The true loading is likely to occur on the millisecond timescale, not the nanosecond to low microsecond timescale. That said, I don’t mean to detract from the result in Figure 2, as this is likely quite solid in my opinion given the nearly 1 microsecond simulations and the replicates showing the same results.

Thank you for your valuable suggestions. It is important to note that we calculated different physical quantities compared to those reported in Argudo’s study. In Figure 3b, we calculated the torque ( instead of the energy, although they share the same dimensional units) of the long AR bundle (AR9-29 of the four filaments combined) and subsequently determined its torsion coefficient. Argudo’s study calculated the torsional spring constant (𝑘ɵ) of three 6-AR-unit stretches of one filament, which were designated as ANK1 (AR 12-17), ANK2 (AR 17-22) and ANK3 (AR 22–27). As the four filaments are coupled within the bundled structure and the torsional axes differ between an individual filament and the four-filament bundle, a direct comparison of the torsional spring constants reported in the two studies is not meaningful.

We agree that extending the simulation time may provide deeper insights into the convergence of energy values. In accordance with your suggestion, we conducted additional simulations to further investigate convergence and compare the results with our existing data, thereby ensuring robustness and consistency. Specifically, we slowed down the original operation of twisting from 10 degrees over 100 ns to 10 degrees over 200 ns, and extended the holding time for selected frames (sampled every 2.5 degrees) from 100 ns to 200 ns. We have updated Figure 3 and relevant main text accordingly (page 7). The results of the new simulations are similar to those of the previous ones, with the fitted torsion coefficient revised from (2.31 ± 0.44) × 103kJ mol−1 ra−1 1 to (2.30 ± 0.31) × 103 kJmol−1 rad−1 This close agreement indicates that our simulations are well-converged. Additionally, we updated the compression–twist coupling coefficient, , from (1.67 ± 0.14) nmrad−1 to (1.32 ± 0.11) nmrad−1

As you suggested, we conducted an additioanl analysis to determine whether the loading velocity/force with the steered MD is sufficiently slow to facilitate the relaxation of the protein and its adoption of lower-energy configurations during the transition. For simulations involving the application of membrane-normal or membrane-parallel force on the TRP domain, we utilized DSSP (Define Secondary Structure of Proteins) analysis to assess the stability of the secondary structure of the TRP domain. The results indicated that, during the application of external forces, the secondary structure of the TRP domain maintained good stability, as illustrated in Figure S11. For simulations involving the rotation of the AR domain, we also analyzed the DSSP of the AR9 to AR11 units, which are positioned directly above the AR8 domain where the twisting force is applied. The secondary structure of the AR domain also exhibited good stability (Figure S12). These are briefly discussed in the Methods section of the revised manuscript (page 17).

It is unclear to me that the force transmission analysis in Figure 4 provides much insight into the mechanics of opening. Perhaps the argument was made, but I did not appreciate it. Related to this the authors state that the transfer velocity is 1.8 nm/ps based on their previous study. Is this value profound or is it simply the velocity of sound in the protein?

The analysis of force transmission presented in Figure 4 offers detailed insights into the transfer of force along the AR domain. While this may appear straightforward, the information elucidates how a pushing force can induce a twisting force during its transmission through the AR spring structure, as well as the primary contributions that stabilize this transmission pathway. To enhance clarity, we have included an additional discussion on page 9.

The force transfer velocity is expected to align with the velocity of sound within the protein. The value of 1.8 nm/ps, however, is specific to the unique structure of the AR spring, which is quite interesting to report in our opinion. Additionally, this rapid transfer speed suggests that the simulation timescale is sufficient for enabling the transfer of compression force from the bottom of the AR domain to the TRP domain in our simulations, given that the simulation timescale is considerably longer than the force propagation timescale within the protein.

The methods description is largely complete, but is missing some details on the MD simulations (barostat, thermostat, piston constants, etc.).

Thank you for pointing out the missing details; we have added the additional information in the revised Methods section.

References

Nadezhdin, K. D., A. Neuberger, Y. A. Nikolaev, L. A. Murphy, E. O. Gracheva, S. N. Bagriantsev, and A. I. Sobolevsky (2021). Extracellular cap domain is an essential component of the trpv1 gating mechanism. Nature communications 12(1), 2154.

Wang, Y., Y. Guo, G. Li, C. Liu, L. Wang, A. Zhang, Z. Yan, and C. Song (2021). The pushto-open mechanism of the tethered mechanosensitive ion channel nompc. Elife 10, e58388.

Version 2

Abstract

Abstract

NOMPC, a tethered mechanosensitive ion channel belonging to the transient receptor potential (TRP) family, converts mechanical stimuli into ionic electric signals that excite neuronal cells. Previous investigations have demonstrated that a pushing force applied to the linker helix domain or the compression of NOMPC’s ankyrin repeat (AR) domain can trigger channel opening. In this study, we explored the direct mechanical causes of NOMPC channel opening as well as the torsional properties of the AR domain, using all-atom molecular dynamics simulations. Our results indicate that a torque directed toward the extracellular side, exerted on the amphipathic TRP domain, is the primary factor driving channel opening. The coupling between compression and twisting of the AR domain ensures that both types of deformation can induce channel opening. Therefore, we propose a twist-to-open model, facilitated by the compression-twist coupling property of the AR domain, to provide further insight into the gating mechanism of the NOMPC channel.

Introduction

Mechanosensitive (MS) ion channels are critical components of mechanosensory transduction, playing an essential role in the perception of sensory stimuli in living organisms. In both vertebrates and invertebrates, mechanosensitive ion channels are widely expressed in peripheral sensory neurons located near or within the surface tissues responsible for detecting mechanical stimuli. Their primary function is to receive and transduce mechanical signals from the external environment, ultimately converting these signals into electrical impulses. This intricate process, known as mechano-electrical transduction (MET), induces changes in neuronal excitability, thereby facilitating the transmission of vital information to higher-order neural centers. 1

Among these, NOMPC, a transient receptor potential (TRP) superfamily MS channel identified in C. elegans, Drosophila, and zebrafish, plays a central role in various mechanosensing-related activities, including soft touch perception, auditory function, and locomotion. 2–6 A salient characteristic that distinguishes NOMPC from its counterparts is its unique intracellular domain, which comprises twenty-nine ankyrin repeats arranged into a distinctive spring-like conformation. This exceptional architecture, proposed to be associated with microtubules within cellular contexts, designates NOMPC as the only known standalone tether-gating MS channel. 7 Consequently, in contrast to the extensively studied force-from-lipid MS ion channels, 8,9 a detailed investigation into the specific molecular mechanism of NOMPC gating is poised to provide fundamental insights into the field of tethered MS ion channels.

In previous studies, it was proposed that compression of the intracellular ankyrin repeat (AR) domain, or a pushing force exerted on the AR domain from the intracellular side, could facilitate the opening of the channel. 10,11 Interestingly, the TRP domain, widely recognized as crucial in the gating process of TRP channels, 12 was observed to tilt up when viewed along the membrane surface, and rotate clockwise when viewed from the intracellular side. 11 It appears that compression of the AR domain can generate both a membrane-normal pushing force and a membrane-parallel twisting force on the TRP domain. 10,11 However, it remains unclear whether the pushing force or the twisting torque acting on the TRP domain is the direct mechanical cause of channel gating. In this work, we determine the critical force component for the gating of NOMPC using all-atom molecular dynamics (MD) simulations. Our results suggest that the twisting force is the key factor in channel gating, while the compression-twist coupling of the AR domain ensures that the compression of the AR domain can also activate the channel.

Results

A membrane-parallel torsion force was observed on the TRP domain during the push-to-open gating process

For a comprehensive exploration of the atomistic gating details through molecular dynamics (MD) simulations with an affordable computational cost, we employed a divide-and-conquer strategy as depicted in Fig. 1(a). Essentially, the membrane-protein simulation system was divided into two subsystems. System I was composed of the transmembrane (TM) region, intracellular linker helices (LH) domain, and ankyrin repeat 29 (AR29). The TM region contains six transmembrane helices (S1-S6) in each chain. Four S6 helices of each chain collectively constitute the central pore of NOMPC, and the TRP domain is sandwiched between the TM domain and the LH domain. System II included the LH domain and twenty-nine ankyrin repeat (AR) units. The four spring-like AR chains of the tetrameric protein were organized into a bundled structure, which is believed to be a crucial element in mechanical perception. 13–15

Force distribution during the push-to-open gating process of NOMPC.

(a) The simulation systems of NOMPC. The NOMPC molecule was divided into two subsystems, denoted by the orange and purple rectangular boxes, for molecular dynamics (MD) simulations. The interface between the linker helices (LH) and the TRP domains is shown in a circular inset (bottom view). (b) The membrane-parallel (purple line) and membrane-normal (yellow line) components of the net forces on the TRP domain exerted by the LH domain. The error bars denote the standard deviations from two compressive trajectories (CI-1/2) and two free trajectories (FI-1/2). (c) The net forces on the TRP domain exerted by the LH domain. The grey quiver on each residue showed the net force exerted by the LH domain. The black showed the resultant force on the residues of the TRP domain. Two major points of force application, E1571 and R1581, were marked. ⊗ represented the inward force component. The data were averaged for the four subunits by rotationally symmetrized around the protein center.

Building upon the previously proposed push-to-open model, 11 we performed steered molecular dynamics (SMD) simulations for system I. In this context, a pushing/compressive force was applied to the ankyrin repeat 29 (AR29), pointing from the intracellular to the extracellular side (trajectories: CI-1/2). Consistent with our previous findings, the transmembrane central pore of NOMPC exhibited a discernible dilation, accompanied by the upward tilt and clockwise rotation of the TRP domain (Fig. S1). The upward tilt was probably caused by the membrane-normal component of the applied mechanical force, while the clockwise rotation was generated by the membrane-parallel component that led to a torque pointing to the extracellular side, raising the question that which of these two components was the direct driving force for the channel opening.

To address this pivotal question, we quantitatively analyzed the force distribution in the above simulations CI-1/2, so that we could decompose the force into the membrane-normal and membrane-parallel components to investigate which one could open the channel. In particular, we used the force distribution analysis (FDA) method 16 to quantify the interactions between pairwise residues along the force transmission pathway within system I (Fig. 1(a)). According to our simulation condition and trajectory analyses, the force transmission initiated from the AR29 domain, traversed through the LH domain, and then proceeded to influence the geometrically adjacent yet non-covalently bonded TRP domain, ultimately guiding the S6 domain and instigating the pore opening. Our analysis showed that the strongest interactions between these domains were predominantly located between a few pairs of residues at the interfaces (Fig. S2), including A1118-N1148, Y1109-D1142 on the AR29-LH interface, and K1244-E1571, D1236-R1581, L1256-S1577 on the LH-TRP interface. Their proximity in space made their Coulomb and Lennard-Jones interactions stronger upon the conformational changes. This facilitated the stability of hydrogen bonds, particularly those between K1244-E1571, D1236-R1581, and L1256-S1577, with occupancies of 57%, 71%, and 24%, respectively. This finding was consistent with the hydrogen bond pairs identified in previous studies when a pushing force was applied. 11

To provide a clear depiction of the net mechanical force distribution resulting from the externally applied mechanical force, we subtracted the FDA results of the force-free simulation trajectories (trajectory: FI-1/2) from the simulation trajectories with pushing/compressive force (trajectory: CI-1/2). The residual force distribution was defined as the net-FDA results (Fig. S3). By analyzing the net mechanical force distribution between the LH and TRP domains using the net-FDA method, we observed that, in addition to the membrane-normal force component, the total net mechanical force transmitted from the LH domain to the TRP domain also exhibited a membrane-parallel torsion force component. This generated a torque directed toward the extracellular side, perpendicular to the membrane (Fig. 1(b)-(c)). Moreover, the transmitted force was primarily concentrated on the E1571 and R1581 residues of the TRP domain (Fig. 1(c)).

To determine the role of the TRP helix in gating, we analyzed the force transmission from the TRP domain to the transmembrane region. The FDA results indicated that the TRP domain interacted with four transmembrane segments, including portions of the S1 helix, the S2-S3 linker, the S4-S5 linker, and S6 (Fig. S4(a),(c),(d)). The strongest interaction occurred with a segment of the S6 domain that encompasses the constriction site of the transmembrane pore (Fig. S4(a)). Consequently, the TRP domain appears to exert an outward force on S6, potentially contributing to gating (Fig. S4(b)). Furthermore, the TRP domain interacts strongly with D1360, located on the S2-S3 linker (Fig. S4(a)), and with S1421 and F1422, located on the S4-S5 linker (Fig. S4(a)). The presence of two pairs of hydrogen bonds between D1360 (S2-S3 linker) and K1578 (TRP), as well as S1421 (S4-S5 linker) and W1572 (TRP) was observed in our simulation and has also been reported in previous work. 7,11

The FDA results at the interface between the transmembrane and intracellular domains elucidated a discernible pathway of force transfer, primarily through several residues located at the AR-LH-TRP interface. In the TRP domain, both membrane-normal and membrane-parallel force components were observed, providing a foundation for further investigation into which component serves as the primary factor in channel gating. As previous studies showed that a clockwise rotation of the TRP domain may be related to channel opening, it was of particular interest to study the effect of the torsion force on the channel structure.

The membrane-parallel torsion force can open NOMPC, but not the membrane-normal pushing force on the TRP domain

Motivated by the above net-FDA results, we aimed to separate the contributions of the membrane-normal and membrane-parallel force components to the gating process and pinpoint whether one or both of these components played an indispensable role in gating. To achieve this, we needed to determine the point of application, the magnitude, and the direction of the exerted mechanical force. As the TRP domain is sandwiched between the AR and TM domains and its conformational change was proposed to be directly related to the channel gating, it is a natural choice to select the TRP domain to apply the decomposed forces in the MD simulations.

Notably, our FDA in the last section showed that the predominant forces propagated from the LH domain to the TRP domain were distributed on two key residues, E1571 and R1581 (Fig. 1(b)-(c)). Therefore, we devised simulations to individually apply membrane-normal and membrane-parallel forces on these two residues of the TRP domain, and the magnitude and direction of the exerted forces were equivalent to those derived from the aforementioned net-FDA results (Fig. 1(b)-(c), Fig. 2(a)-(b)). As the force transmission occurred mainly via side-chain interactions along the interface between the LH and TRP domains, we selected the side chains of the E1571 and R1581 as the point of application of the exerted forces.

A membrane-parallel torsion force can open NOMPC.

(a-b) The side view (a) and bottom view (b) of the net-FDA of the E1571 and R1581 on the TRP domain exerted from the LH domain, which was decomposed into the membrane-parallel and membrane-normal components. ⊗ represented membrane-normal forces pointing into the membrane. (c-d) The transmembrane pore size evolution in the simulations with membrane-parallel forces (c) or membrane-normal forces (d), three replicates for each condition (trajectories: MPI-1/2/3, MNI-1/2/3, see Methods).

Our simulations revealed that the membrane-parallel torsion force alone was sufficient to open the gate in all three MD simulation replicates (trajectories: MPI-1/2/3, Fig. 2(c)), whereas the membrane-normal pushing component did not trigger pore opening within the simulation time (trajectories: MNI-1/2/3, Fig. 2(d)). According to the pore size evolution in the trajectories MPI-1/2/3 (Fig. 2(c)), it became evident that the most constricted region along the pore, the gate situated around the hydrophobic I1554, exhibited an initial radius measuring less than 1.0 Å, as calculated by the HOLE program. 17 This configuration represented a closed architecture. Under the influence of the membrane-parallel torsion force, the radius of the lower constriction was significantly dilated to above 2.0 Å after 600 ns. This dilation meant that water molecules could smoothly pass through the channel (Fig. S5), a phenomenon often indicative of the first step of channel opening. This is highly similar to our previous observation when a pushing force was applied to the AR domain. 11

In contrast, in the trajectories MN-1/2/3 with the membrane-normal force on the TRP domain, no pore dilation was observed (Fig. 2(d)) within our simulation time. In fact, the channel appeared to become even more constricted at the upper constriction region (filter), as depicted in Fig. 2(d). This indicated that the membrane-normal force component on the TRP domain was probably not able to open the channel, at least no such sign was observed within our simulation time.

It should be noted that, according to our analysis, the membrane-normal pushing forces exerted on residues 1571 and 1581 are approximately one-third and two-thirds of the membrane-parallel twisting force, respectively. This observation indicates that the twisting component is inherently stronger than the pushing component when a compressive force is applied to the AR domain. To examine the impact of the two force components when they are of equal magnitudes, we conducted additional simulations (Fig. S6, see Methods). To ensure the stability of the protein and membrane structure, we opted to reduce the membrane-parallel force rather than increase the membrane-normal force. As illustrated in Fig. S6, halving the membrane-parallel force resulted in channel opening in two out of three trajectories, while reducing it to one-third resulted in channel opening in one out of three trajectories. These results suggest that, when force components of equal magnitude are applied to the key residues on the TRP domain, the membrane-parallel (twisting) force is more effective in promoting channel opening than the membrane-normal (pushing) force.

Thus, our investigations substantiated that a torque on the TRP domain, pointing to the extracellular side, was the key to driving the channel open. On the basis of the original push-to-open or compress-to-open model, here we provide further details on the force analysis and channel opening, arguing that it is the membrane-parallel torsion force that acts as the direct mechanical cause to open the channel. It should be noted though, this does not mean the push-to-open or compress-to-open model was wrong, as our simulations showed that a pushing force on the LH domain can generate a torsion force on the TRP domain, 11 so does the compression of the AR domain. 10 Apparently, the unique structure of NOMPC ensures that a straightforward pushing force or compression of the AR spring can spontaneously generate a torsion force to open the channel. Therefore, the torsional mechanical properties of the AR domain became more intriguing.

The torsional mechanical properties of the AR domain and force transmission pathway

To study the torsion property of the AR spring, we performed all-atom steered molecular dynamics (SMD) simulations for system II, in which we enforced a gradual clockwise rotation to the lower end of the AR domain (AR1-AR8) at a rate of 0.05 ° ns−1 along the membrane-normal axis of NOMPC while keeping the upper end (LH domain) position-restrained (Fig. 3(a), trajectory: RII). From the trajectory, we extracted structural snapshots at an interval of 2.5° during the twisting process and relaxed these structures for 200 ns with both ends position-restrained. This protocol was employed to further relax the twisted AR spring to improve the accuracy of the calculated torsional mechanical parameters (Fig. S7). Additionally, to ensure the convergence of the RII trajectory calculations, we simulated a set of trajectories with a faster application of the torsional force (trajectory: RII-fast). The results were highly consistent with those of RII (Fig. S8). Please refer to the Methods section for more details.

Torsional mechanical properties of the AR bundle.

(a) The sketch of the steered molecular dynamics simulation. The dotted arrows revealed the direction of twisting. (b) The relationships between the torque on AR1-8 (left Y-axis), the length of the AR domain (right Y-axis), and the rotational angle of AR1-8. This was calculated from the rotating trajectories RII (0-4).

With the above simulation data (trajectory: RII), we calculated the torsional coefficient of the AR bundle, using the equation: where M represented the torque applied to both ends, c denoted the torsion coefficient of the AR domain bundle, and φ was the average torsion angle of AR1-8 with respect to its initial position. As shown by the green plots in Fig. 3(b), the fitted torsion coefficient turned out to be (2.30 ± 0.31) × 103 kJ mol−1 rad−1. Such a torsion coefficient roughly corresponds to a similarly-sized cylindrical volume of uniform polyethylene or rubber material. 18

Interestingly, we observed a gradual increase in the compressive deformation of the AR spring as the torsional angle increased in our simulations (Fig. 3(b)). This implies a coupling between the twist and compression of the AR domain. To quantitatively characterize this coupling property, we defined the compression-twist coupling coefficient . This coefficient represents the compressive deformation induced by a unit torsional deformation. For the AR8-AR29 of NOMPC, the fitted value of the compression-twist coupling coefficient kct was (1.32 ± 0.11) nm rad−1 within a 10-degree torsion range. Indeed, it was also noticed that a torque could be generated during the application of compressive forces in previous investigations. 10,11 In light of these findings, we believe that the unique structure of the AR and LH domains enhances NOMPC’s robustness in sensing external stimuli. Specifically, the compression-twist coupling property of the AR bundle ensures that both the twist and compression of the intracellular domain can generate torque to rotate the TRP domain, thereby facilitating the opening of the pore.

To pinpoint the force transmission pathway along the AR spring while it is twisted, we utilized the same net-FDA method as above to analyze the force distribution during the twisting process of the trajectory RII. The net-FDA results of each (i + 1)th AR unit exerted from the adjacent ith AR unit unveiled a spiral force transmission route, rooted in the helical structure inherent to the AR spring (Fig. 4(a)). Strong pairwise interactions predominantly resided in bonded residue pairs, while only a small fraction was associated with non-bonded residue pairs (Fig. 4(b)). Clearly, these bonded residue pairs, which are integral to the spiral backbone of the AR spring, were critical for transferring external forces. The non-bonded residue pairs were primarily situated at the adjacent turns within the AR α-helices and the interfaces between AR units. To clearly illustrate the details, we have magnified the local structure of AR13-15 (Fig. S9). As depicted in Fig. 4(c)-left, a representative non-bonded net-FDA of the local region of AR13-15 was shown. The data around the diagonal line represented the pairwise residues located at the adjacent turns between α-helices, while the wing-shaped data positioned further away from the diagonal indicated pairwise residues situated at the interfaces between AR units. Apparently, although the bonded residue pair interactions are predominant, the non-bonded interactions, both within the same AR unit and across different AR units, are important for the force transmission too.

Force transmission along the AR domain.

(a) The quiver plot of net-FDA (XY planar) of each (i + 1)th AR unit exerted from the ith AR unit showed the force transmission from the intra-cellular terminal to the transmembrane side of the AR domain viewing from AR1 toward AR29. ⊙ represented the outward Z-component force. (b) The distribution of the pairwise residue-residue net-FDA within AR9 to AR29 (weak interaction under 5 pN neglected). The inset showed local information over 70 pN. (c) The representative interaction heatmaps (AR13-AR15) showed the magnitude of non-bonded pairwise residue-residue net-FDA (left) and hydrogen bonds occupancy (right). (d) The Pearson correlation coefficient (PCC) between net-FDA and hbonds in AR units. Inset showed the mean value and standard deviation of Intra-AR (diagonal) and Inter-AR (sub-diagonal) PCCs.

Furthermore, we noticed that there is a correlation between the distribution of non-bonded net-FDA and the existence/strength of hydrogen bonds, as demonstrated by similar patterns in Fig. 4(c). For clarity, we divided these two 737 × 737 matrices (ResID of AR9-AR29: 400-1136) into 21 × 21 subgroups (grouped by AR unit). Then, we calculated the Pearson correlation coefficient (PCC) between the non-bonded net-FDA and the occupancy of hydrogen bonds of each subgroup, resulting in a 21 × 21 PCC matrix (Fig. 4(d)). Relatively high PCCs in the sub-diagonal grids of Fig. 4(d) (inter-AR), and even higher PCCs in the diagonal grids (intra-AR), are observed (inset of Fig. 4(d)). This underscored the crucial role of the hydrogen bonding network in facilitating force transmission.

Therefore, we conclude that the force transmission route primarily comprises bonded residues that constitute the spiral backbone of the AR chains, supplemented by the hydrogen bonding network formed by adjacent residues within the same or neighboring AR units. Although this may initially seem trivial, it offers valuable insights into the mechanism by which a twisting force can be formed and transmitted along the AR spring architecture. Furthermore, it underscores the stabilizing roles of both covalent and non-covalent interactions in preserving the stability of the force transmission pathway.

Discussion

The unique structure of NOMPC and its fascinating AR spring have sparked great curiosity. Building upon the push-to-open and compress-to-open mechanism, here we propose a twist-to-open model, as illustrated in Fig. 5. At the core of this model was the idea that a twisting force on the TRP domain is the key to the opening of the pore. This is not to say the compress-to-open or push-to-open models are wrong, but that a compression/pushing on the AR domain will lead to a twist force on the TRP domain, and this twist force is the direct cause of the pore opening. The spiral structure of each AR chain, as well as the hydrogen bonding network formed among spatially adjacent residues, enables the AR bundle to convert the compression deformation to a twist deformation, thus exerting a membrane-parallel torsional force on the LH domain. In addition, the structure around the LH domain can also convert a membrane-normal pushing force to a membrane-parallel torsional force, as shown in our previous study. 11 Consequently, the compression of the AR domain, or a pushing force from the intracellular side, can be converted to a torque on the TRP domain, which points to the extracellular side. We believe that this torque is the direct mechanical cause that drives the TRP domain to rotate clockwise (looking from the intracellular side) to open the channel.

A schematic diagram of the twist-to-open model.

The side view (left) and bottom view (right) of the twist-to-open model. Both the compression and clockwise twisting of the AR domain can generate a membrane-parallel twisting force on the TRP domain, which is the key component for gating.

The highly conserved TRP domain serves as a structural characteristic of the TRP protein super-family. In our simulations of NOMPC, we observed a clockwise rotation of the TRP domain during the gating process of the proteins. Interestingly, such rotation was also observed in many other TRP channels. We investigated the likely conformational changes of multiple channels within the TRP family upon opening and found that the rotation of the TRP domain is a universal conformational change, whereas the tilt of the TRP domain is considerably less consistent (Fig. 6). Based on these findings, we believe that the rotation of the TRP domain may be a common feature in the gating mechanism of these TRP channels, and the unique structure and compression-twist coupling property of the intracellular domain of NOMPC ensure its robustness in the perception of multiple forms of mechanical stimuli: not only twist, but also compression, can result in the clockwise rotation of the TRP domain, thereby facilitating channel opening.

Conformational changes of the TRP domain in various TRP channels with resolved closed- and open-state cryo-EM structures.

(a) The membrane-parallel rotational angle of the TRP helix (a positive value indicates clockwise rotation when viewed from the intracellular side). (b) The tilt angle of the TRP domain relative to the membrane surface (a positive value indicates tilt-up). S1-S4 helices were aligned for analysis. (c) and (d) represent the rotational and tilting conformational changes of the TRP domain upon gating, respectively. In the majority of the studied cases, the TRP domain rotates clockwise (c). The exception of counterclockwise rotation is observed for HsTRPA1. (please refer to Table 1 for the protein structures used in this analysis)

Given the universal conformational change of the TRP domain and the preceding analysis of the interactions between the TRP helix and the transmembrane region, it is pertinent to discuss the potential role of this amphipathic helix. As initially suggested in the MS channel MscL, researchers have identified that the amphipathic N-terminus of MscL provides a structural framework for bilayer-mediated gating of MS channels. 19 Subsequent studies have further demonstrated that many MS channels share a common structural feature: amphipathic helices that are directly or indirectly connected to pore-lining regions. 8 These amphipathic helices interact with surrounding lipids and transmit membrane tension, which is consistent with the force-from-lipid gating mechanism. Although the present study primarily focuses on the force-from-tether model, it appears that the amphipathic TRP helix, which is connected to the pore-lining S6, is also crucial to the gating of NOMPC. The observed interaction between the TRP domain and the S4–S5 linker suggests that these two structural elements may function together as a composite amphipathic unit, potentially playing a significant role in perceiving and propagating gating signals. In addition, the TRP helix and the S4-S5 linker form the binding site for ligand-gated TRP channels, such as TRPV1. Therefore, we propose that the combined TRP-S4-S5 unit may constitute a key structural element responsible for multimodal gating mechanisms within the TRP channel family.

Mammalian TRPs possess shorter ankyrin repeats than NOMPC, raising the question of whether they can also convey force in a manner similar to NOMPC. As an initial exploration, we selected the full-length squirrel TRPV1 (PDB: 7LQY) 20 and applied pushing or torsional forces to the intracellular end of TRPV1, separately (Fig. S10(a)). Rotation of the TRP domain was observed under both types of mechanical stimulation (Fig. S10(b-e)). The torsional force-induced conformational change in the TRP domain resembles that observed in NOMPC. This suggests that a torsional force applied to the shorter ankyrin repeats of mammalian TRPs may generate similar effects on channel gating. However, since these ARs are not tethered, the implications of these results in the context of biological functions remain unclear.

Although we calculated the detailed interactions between residue pairs of NOMPC, it should be noted that, due to the limitations of the FDA method, our force analysis lacked the entropic force, which may also play a significant role in the force transmission along the AR spring. Nonetheless, this won’t affect our twist-to-open model, since the entropic effect was present in our MD simulations. Another limitation of the present study is that the simulation time is very short compared to the physiological timescale, so we only observed a partially open state of the structure in our simulations and cannot exclude the possibility that other factors, such as the membrane-normal force component, may also assist in the gating process. However, we found that the mechanical stimuli are transferred at a fast speed along AR springs in simulations, at around 1.8 nm ps−1. 11 Therefore, the mechanical stimuli were nearly instant for our simulation time, which allowed us to determine the trend or the direction of the conformational changes, even though the observation of the complete gating process is still challenging. In this context, we think that the dilation of the pore under a torsion force on the TRP domain is a piece of strong evidence that twist is the key to the channel opening.

Methods

The all-atom simulation Systems

To set up the molecular dynamics (MD) simulations, we first prepared the simulation systems using the cryo-EM structure of NOMPC (PDB: 5VKQ). 7 CHARMM-GUI was used to generate the configuration and topology of the simulation systems. 43 The parameter files were created for the CHARMM36m force field. 44,45

The system I included the TM domain, the LH domain, and the AR29 of NOMPC. The PPM server was used to reorient the NOMPC structure to ensure that the TM domain of NOMPC was well located in a lipid bilayer. 46 The protein was embedded in a POPC bilayer. Then, we solvated the protein-membrane complex in a water box using the TIP3P water model. Counterions were added to neutralize the system, ensuring physiological ionic strength by adding Na+ and Cl− ions to achieve a concentration of 0.15 M. The details of simulation system I are shown in Table 2.

The system II included the LH domain and the AR domain (AR1-AR29) of NOMPC. The protein structures were immersed in a rectangular box of TIP3 water molecules and 0.15 M NaCl. The details of simulation system II are shown in Table 2.

All-atom MD simulations

All of the MD simulations were performed with GROMACS 5.1.2. 47 The REDUCE program in AMBER was used to add hydrogens to the original PDB files and determine the protonation state of the histidine residues. 48,49

For system I, we used the steepest descent algorithm to achieve energy minimization. Then, following up with a two-stage equilibration, a 0.4 ns NVT equilibration simulation with harmonic restraint was applied to the protein molecule (force constants of 4000 kJ mol−1 nm−2 on the backbone and 2000 kJ mol−1 nm−2 on the side chains), and a 20 ns NPT equilibration simulation with gradually decreased restraint (from 2000 to 100 kJ mol−1 nm−2 on the backbone and from 1000 to 50 kJ mol−1 nm−2 on the side chains). During the equilibration processes, planar restraints were used to keep the positions of lipid head groups along the membrane-normal direction. The simulation temperature of the system I was set to 300 K. The time step was 2 fs. The cubic periodic boundary condition was used during the simulations, and the van der Waals interaction was switched off from 10 to 12 Å. The long-range electrostatic interactions were calculated using the particle mesh Ewald (PME) method. 50 The Nosé–Hoover thermostat 51 was applied separately to the protein, membrane and solvent with a coupling constant of 1.0 ps and reference temperature of 300 K. Semi-isotropic pressure coupling was applied using the Parrinello–Rahman algorithm 52 with a reference pressure of 1 bar and a coupling constant of 5.0 ps. The compressibility was set to 4.5 × 10−5 bar−1. All bonds involving hydrogen were constrained using the LINCS algorithm. 53

For system II, we used the steepest descent algorithm to achieve energy minimization. Then, also followed by a two-stage equilibration, a 0.2 ns NVT equilibration simulation with harmonic restraint was applied to the protein molecule (force constants of 400 kJ mol−1 nm−2 on the backbone and 40 kJ mol−1 nm−2 on the side chains), and a 10 ns NPT equilibration simulation with harmonic restraint (force constants of 400 kJ mol−1 nm−2 on the backbone and 40 kJ mol−1 nm−2 on the side chains). The simulation temperature of the system II was set to 300 K. The time step was 2 fs. The cubic periodic boundary condition was used during the simulations, and the van der Waals interaction was switched off from 10 to 12 Å. The long-range electrostatic interactions were calculated using the particle mesh Ewald (PME) method. 50 The Nosé–Hoover thermostat 51 was applied separately to the protein and solvent with a coupling constant of 1.0 ps and reference temperature of 300 K. Semiisotropic pressure coupling was applied using the Parrinello–Rahman algorithm 52 with a reference pressure of 1 bar and a coupling constant of 5.0 ps. The compressibility was set to 4.5 × 10−5 bar−1. All bonds involving hydrogen were constrained using the LINCS algorithm. 53

Steered MD simulations

Steered molecular dynamics (SMD) simulations were utilized to probe the mechanical gating mechanisms of the ion channels. SMD was performed by attaching a virtual spring to a specific group of atoms. To gradually apply a force to the designated group of atoms and prevent sudden conformational changes that could cause structural damage, a constant velocity was first applied to the spring to pull the molecule along a predefined reaction coordinate, followed by the application of a constant force. The force exerted on the molecule was recorded throughout the simulation. To ensure reproducibility, multiple independent SMD simulations were performed in this research.

The SMD simulation detail of the compressive trajectories of system I (CI-1/2) and free trajectories of system I (FI-1/2) can be found in previous work. 11

For trajectories MPI-1/2/3 (membrane-parallel force component on system I Replicate-1/2/3) and MNI-1/2/3 (membrane-normal force component on system I Replicate-1/2/3) of system I (Fig. 2(c)-(d)), we separately exerted membrane-parallel (MPI-1/2/3) or membrane-normal (MNI-1/2/3) external force on the side chain of E1571 and R1581 (reference group: S1-S4), as shown in Table 3. At the first 200 ns, we applied an increasing force by using the constant-rate pulling method. The pulling was performed with a spring constant of 100 kJ mol−1 nm−2 and a pulling velocity of 0.005 nm s−1. Then we applied the calculated constant force (Table 3) by using the constant-force pulling method in the last 600 ns. For additional simulations with smaller magnitude of membrane-parallel force component, we have trajectories MPI-h1/2/3 (half of membrane-parallel force component on system I Replicate-1/2/3) and trajectories MPI-t1/2/3 (one third of membrane-parallel force component on system I Replicate-1/2/3), as shown in Table 3.

Besides, to prevent the protein from rotation as a whole, we used the enforced rotation method, 54 setting the rotation speed as zero for the S1-S4 helices in the transmembrane region of the protein. The rotation potential was rm2 with a force constant of 25 kJ mol−1 nm−2 and the pivot point for the potential was the center of the protein. The norm method was used to determine angle of rotation group. Distance restraints were applied between the acceptors and donors on the side chains of residue pairs 1564-1568 and 1567-1571, to maintain the hydrogen bonds and prevent structural damage in the α-helix under the applied mechanical force.

The SMD simulation temperature of the system I was set to 300 K. The time step was 2 fs. The cubic periodic boundary condition was used during the simulations, and the van der Waals interaction was switched off from 10 to 12 Å. The long-range electrostatic interactions were calculated using the particle mesh Ewald (PME) method. 50 The Nosé–Hoover thermostat 51 was applied separately to the protein, membrane and solvent with a coupling constant of 1.0 ps and reference temperature of 300 K. Semi-isotropic pressure coupling was applied using the Parrinello–Rahman algorithm 52 with a reference pressure of 1 bar and a coupling constant of 5.0 ps. The compressibility was set to 4.5 × 10−5 bar−1. All bonds involving hydrogen were constrained using the LINCS algorithm. 53

For trajectory RII (Rotation of system II), we performed a three-step simulation to better study the torsional mechanical properties of the AR domain, as shown in Table 4.

In the first step, we aimed to relax the structure by fixing the relative rotational angle between both ends along the rotational axis, which is the Z axis of system II. We choose to manipulate AR1-AR8 as a whole, considering the fact that the cryo-EM structure of AR1-AR7 was not resolved but modeled. Therefore, it was prudent to exert and transmit torque starting from the more reliably resolved structure, i.e., from AR8 upward, while preserving the integrity of the entire protein structure. With the help of the enforced rotation function of GROMACS, 54 we used the pm rotational potential on the bottom terminal of the AR domain (AR1-AR8) with a force constant of 100 kJ mol−1 nm−2 and a zero angular velocity for 20 ns, and the up terminal (LH domain) with a force constant of 50 kJ mol−1 nm−2 and a zero angular velocity for 20 ns. The pm rotational potential can constrain the group of atoms to rotate within the rotational plane as required, while allowing free movement in the direction perpendicular to the rotational plane. This means that the entire AR bundle will be subjected to torsion, while being able to stretch and compress freely. The pivot point for the potential was the center of the protein. The norm method was used to determine the angle of the rotation group.

In the second step, we clockwise rotated the AR1-8 along the Z axis at an angular velocity of 0.05 ° ns−1 with a force constant of 100 kJ mol−1 nm−2 and, while keeping the LH domain still (Fig. 3(a)) with pm rotational potential with a force constant of 50 kJ mol−1 nm−2. The difference in rotation angle between the bottom and up terminal gradually increased to 10° over 200 ns (Fig. S7, trajectory: RII-0). The pivot point for the potential was the center of the protein. The norm method was used to determine the angle of the rotation group.

In the third step, when rotating at every 2.5° in trajectory RII-0, we extracted structural snapshots and held both ends of the AR domain for 200 ns to get a better equilibrated structure (Fig. S7, trajectory: RII-1 for the 2.5°-rotational angle, RII-2 for 5°, RII-3 for 7.5°, and RII-4 for 10°), and the readouts of torques on both ends gradually increased with opposite directions. The other parameters for enforced rotation were kept consistent with those used in the previous two steps. Specifically, we used the trajectory RII-1/2/3/4 to calculate the torsional mechanical properties of the AR bundle and the trajectory RII-0 to study the force transmission pathway along the AR chain. When the torsional angles of the AR bundle are 2.5i° (i = 1, 2, 3, 4), we analyze the 200 ns-long RII-i trajectories. For the torques at both ends of the AR, we collect torque data on the LH domain every 2 ps. Under equilibrium conditions, the torques at the top and bottom of the AR bundle are equal in magnitude but opposite in direction. To assess the length change of the AR bundle, we measure the distance between the centroids of the LH domain and AR8 unit.

To further investigate the convergence of the torsional mechanical properties of the AR bundle and better compare the results with trajectory RII, we conducted additional simulations with a faster velocity of force loading to ensure robustness and consistency. We adopted the same first step as in trajectory RII. The rotational speed in the second step was doubled to 0.1 ° ns−1, resulting in a total twist of 10° within 100 ns. The holding time in the third step was adjusted to 100 ns. This accelerated simulation trajectory is referred to as RII-fast (see Table 4). All other simulation parameters remained unchanged. The mechanical properties derived from the RII and RII-fast trajectories were found to be close (see SI), indicating that the results from the RII trajectory set are converged.

The SMD simulation temperature of the system II was set to 300 K. The time step was 2 fs. The cubic periodic boundary condition was used during the simulations, and the van der Waals interaction was switched off from 10 to 12 Å. The long-range electrostatic interactions were calculated using the particle mesh Ewald (PME) method. 50 The Nosé–Hoover thermostat 51 was applied separately to the protein and solvent with a coupling constant of 1.0 ps and reference temperature of 300 K. Semi-isotropic pressure coupling was applied using the Parrinello–Rahman algorithm 52 with a reference pressure of 1 bar and a coupling constant of 5.0 ps. The compressibility was set to 4.5 × 10−5 bar−1. All bonds involving hydrogen were constrained using the LINCS algorithm. 53

We used DSSP analysis (Define Secondary Structure of Proteins) to ensure the stability of the secondary structure of NOMPC when applying SMD, for the TRP domain in system I and the AR units in system II (Fig. S11, Fig. S12).

Force distribution analysis

Force distribution analysis (FDA) focuses on computational efficiency and low-memory usage and makes possible time series analysis of pairwise forces variation in long MD simulations and for large molecular systems. 16 It can compute atomic pairwise forces for all types of bonded interactions as well as Coulomb and Lennard-Jones potential by rerunning the trajectories. By summing up all the atomic pairwise forces within the two residues of interest, it can also calculate the pairwise force of the interaction between residues ri and rj, which is labelled as . Similarly, if we focus on the interaction between two domains, we can sum up all the pairwise forces of the residues from these two domains.

In the calculation, we made an input file which contained all the parameters that FDA needed to run. We set onepair = detailed and type = all so that all pairwise forces of different types between the same pair of residues are stored separately. The option residuebased was set as pairwise_forces_vector so that the residue-based pairwise forces were written to output files as vector values. The forces for every frame were written by setting time_averages_period = 0.

Then, forces sampled by MD simulations under two conditions (with and without perturbation) were compared. We analyzed the interactions between residues in the trajectories with the external force (CI-1/2), labelled as , and without the external force (FI-1/2), labelled as , respectively. Then the average net force distribution induced by the compressive external force, which was named net-FDA, , was obtained by subtracting the FDA result of trajectories without the external force from the FDA result of trajectories with the external force, In the end, we averaged the vector net-FDA results of the four subunits by rotationally symmetrized around the protein center.